Comparing Apples and Oranges in Youth Depression Treatments?

### 2024-08-14 revision1_reply_to_reviewers.qmd

---
title: "Comparing Apples and Oranges in Youth Depression Treatments? A Quantitative Critique of the Evidence Base and Guidelines"
subtitle: "Major Revision: Letter to Reviewers"
format: docx
editor: visual
---

```{r}
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)
library(broom)
library(effsize)
library(here)

#set source
load("my_environment.RData")
```

@@

The ENcePP [also clearly states that](https://encepp.europa.eu/encepp-toolkit/methodological-guide/chapter-7-effect-modification-and-interaction_en):

The presence of effect modification depends on which measure is used in the study (**absolute or relative** \[added emphasis by us\]**)** and can be measured in two ways: on an additive scale (based on risk differences \[RD\]), or on a multiplicative scale (based on relative risks \[RR\]).

To make this clear, we reproduce the table above, only adding another two column, one is the Odds Ratio (OR) another relative/multiplicative measure of relative risk that the reviewer also refers to as "more generalizable", and that of absolute risk, AR (i.e. the risk difference), an additive measure.

|         |                          |                               |     |     |      |
|------------|------------|-------------|------------|------------|------------|
|         | Control event rate (CER) | Experimental event rate (EER) | RR  | OR  | AR   |
| Trial 1 | 5/100                    | 10/100                        | 2.0 | 2.1 | 0.05 |
| Trial 2 | 30/100                   | 60/100                        | 2.0 | 3.5 | 0.3  |

@@

As can be seen the ORs differ quite substantially and the ARs differ by an order of magnitude. There is clearly an effect modification of some sort going on, when considering the OR or AR. Indeed, as the EMA and ENcePP state, whether effect modification is present (also) depends on the measure used. We therefore disagree with the reviewer that differences in the control event rate are ipso facto prognostic factors (PFs) if they do not affect the RR. Importantly, the RR is notorious for being potentially misleading (<https://pubmed.ncbi.nlm.nih.gov/15823376/>) and the necessity of reporting absolute risks is emphasised in standard clinical publications (e.g. <https://academic.oup.com/ndt/article/32/suppl_2/ii13/3056571> ).

To illustrate this point further in the context of trials, we point out that a key metric in intervention evaluation is the Numbers Needed to Treat (NNT) or its counterpart Numbers Needed to Harm (NNH). The NNT simply defined as the inverse of the AR, namely:

$$
NNT = \frac{1}{AR}
$$

In the case above:

$$
NNT_{Trial_1} = \frac{1}{0.05} = 20 
$$

whereas:

$$
NNT_{Trial_2} = \frac{1}{0.3} = 3.3 
$$

@@

The interpretation of NNT in the context of a placebo-controlled trial is simply: the average number of people who need to be treated with active treatment rather than placebo for one additional person to benefit. It is obvious from the above that, were such numbers to occur in a clinical trial, the clinical and public health implications of the difference between two treatments would be potentially vast and that it would be wrong to have considered RRs alone.

But we would go a step further and argue that all summary measures, including the AR, can conceal important clinical and public health facts, and that therefore reporting data that are close as possible to the raw numbers is key. We will illustrate this with an example.

Let there be two treatments A and B, with data on their efficacy coming comparisons that each has had with A_control and B_control, respectively. The efficacy of A is defined as:

$$
Eff_{A} = A - A_{Control} = 90% - 60% = 30%
$$

and that of B as:

$$
Eff_{B} = B - B_{Control} = 40% - 10% = 30%
$$

And therefore:

$$
Eff_A = Eff_B
$$

@@

#### Response to 3.8

Before we respond to the reviewer's point one by one, we would like to suggest that a more useful---not least because more formally stated and generalizable---framework for theoretical considerations is the issue of external validity. We had included these formal treatments in previous versions of the manuscript, but decided to remove them becauseBriefly,

The average treatment effect is the measure of interest, defined as:

$$
ATE = E\[Y(1)\] - E\[Y(0)\] 
$$

And the conditional treatment effect:

$$
CATE(X = x) = E\[Y(1) - Y(0) \\mid X = x\]
$$
